---
title: "test 1"
author: "Charles Reid"
date: "5/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
# read in libraries 
library(tidyverse) # general dataframe manipulation
library(MASS) # statistical package
library(rms) # regression modeling package
library(prettydoc)
library(lubridate)
library(ggplot2)
library(dplyr)
library(data.table)
library(ggrepel)
library(tidyverse)
library(stargazer)
library(kableExtra)
library(flexdashboard)
library(DataExplorer)
```

```{r include=FALSE}
accepted_loans <- read_csv("LoanData.csv")
```

## Regression

Lending Club provides personal loan, business loan, auto refinancing and patient solutions with APR ranging from 4.99%-35.89% given the differences of the borrowers’ characteristics. The loan amount is between 1,000 dollars to 500,000 dollars with a payment term of 1-5 years. Lending Club uses an intermediate pricing model (Galloway, 2009) which means the interest rate is predetermined by the platform.


```{r}
requested_balance_vs_granted_balance <- lm(funded_amnt ~ loan_amnt, data = accepted_loans)
stargazer(requested_balance_vs_granted_balance, type = "text")
```

```{r}
general_factors_affecting_approved_balance <- lm(funded_amnt ~ annual_inc + emp_length + home_ownership + purpose + delinq_2yrs + term + int_rate + grade, data = accepted_loans)

stargazer(general_factors_affecting_approved_balance, type = "text")
```
```{r}
employment_factors_affecting_balance <- lm(funded_amnt ~ annual_inc + emp_length + home_ownership + purpose + delinq_2yrs + term + int_rate + grade, data = accepted_loans)

stargazer(employment_factors_affecting_balance, type = "text")
```

```{r}
factors_affecting_interest_rate <- lm(int_rate ~ fico_range_low	+ fico_range_high + inq_last_6mths + open_acc + pub_rec + revol_bal	+ revol_util + total_acc	+ collections_12_mths_ex_med + inq_fi + avg_cur_bal, data = accepted_loans)

stargazer(factors_affecting_interest_rate, type = "text")
```

```{r}
#factors_affecting_hardships <- lm(hardship_flag ~ revol_bal + revol_util + total_acc + total_rec_late_fee + collections_12_mths_ex_med + max_bal_bc + inq_fi + tax_liens + fico_range_low + fico_range_high + delinq_amnt + delinq_2yrs + dti, data + pymnt_plan,family = binomial(link = "logit") ,data = accepted_loans)

#stargazer(factors_affecting_hardships, type = "text")
```


